---
title: 'Spain Speaks: Spain''s Battle w/ COVID-19'
author: "David J Jackson"
date: "7/30/2020"
output: 
  html_document: 
    fig_width: 9
    fig_height: 6
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Comparing Spain's COVID-19 Cases and Deaths to Top Five EU Cointries 
* Spain = ES; Germany = DE, France = FR 
* Poland = PL,Britain = UK, Italy = IT


## Import data from European CDPC
Data Source: https://www.ecdc.europa.eu/en/publications-data/download-todays-data-geographic-distribution-covid-19-cases-worldwide

```{r, echo=FALSE}
rm(list=ls())
EUP <-read.csv("../DATA/EU_top_10_pop..csv")
EUPOP <- subset(EUP,Population >=30000000)
EUPOP$Country <- as.factor(EUPOP$Country)
EUCOVID <- read.csv("../DATA/COVID-19-ECDPC-2020-07-31.csv")
EUCOVID$dateRep <- gsub("/","-",EUCOVID$dateRep)
EUCOVID$dateRep <- as.Date(EUCOVID$dateRep,format="%m-%d-%Y")
```
```{r,echo=FALSE}
DE <- subset(EUCOVID,geoId =="DE")
FR <- subset(EUCOVID,geoId =="FR")
UK <- subset(EUCOVID,geoId =="UK")
PL <- subset(EUCOVID,geoId =="PL")
IT <- subset(EUCOVID,geoId =="PL")
spain <- subset(EUCOVID,geoId =="ES")
EUCOMBINED <- rbind(DE,FR,IT,PL,UK)
```

#### Quantile info for Cases:  `r quantile(EUCOMBINED$cases)`
#### Quantile info for Deaths: `r quantile(EUCOMBINED$deaths)`

```{r,echo=FALSE}
barplot(Population~Country,EUPOP, main="Top 5 EU Counties by Population")
#
```

## Calc Daily Sun and Mean for Cases and Deaths for Gang of 5.

```{r, echo=FALSE}
EUTC <- aggregate(x=EUCOMBINED$cases,by=list(EUCOMBINED$dateRep),FUN=sum)
EUMC <- aggregate(x=EUCOMBINED$cases,by=list(EUCOMBINED$dateRep),FUN=mean)
```

```{r, echo=FALSE}
EUTD <- aggregate(x=EUCOMBINED$deaths,by=list(EUCOMBINED$dateRep),FUN=sum)
EUMD <- aggregate(x=EUCOMBINED$deaths,by=list(EUCOMBINED$dateRep),FUN=mean)
```
```{r, echo=FALSE}
plot(EUTC$Group.1,EUTC$x, main="EU Gang of 5 Daily Cases(mean and Sum)",type="l",xlab="Date Reported",ylab="Daily Cases: Mean & Total",col="blue")
lines(EUMC$Group.1,EUMC$x,col="red")
legend("topleft",
       c("Total Cases","Mean Cases"),
       col=c("blue","red"),
       lty = c(1, 1),
       cex = 0.6)
grid()
```
```{r,echo=FALSE}
plot(EUTD$Group.1,EUTD$x, main="EU Gang of 5 Daily Deaths(mean and Sun)",type="l",xlab="Date Reported",ylab="Daily Cases: Mean & Total",col="blue")
lines(EUMD$Group.1,EUMD$x,col="red")
legend("topleft",
       c("Total Deaths","Mean Deaths"),
       col=c("blue","red"),
       lty = c(1, 1),
       cex = 0.6)
grid()
```

## Compare Spain's Daily Cases and Death to Gang  of 5

```{r,echo=FALSE}
plot(EUMC$Group.1,EUMC$x, main="EU Gang of 5 Daily Cases(mean and Sun",type="l",
     xlab="Date Reported",
     ylab="Daily Cases: Mean & Total)",col="red",ylim=c(0,max(EUMC$x)))
lines(spain$dateRep,spain$cases,col="blue")
legend("topleft",
       c("Mean Deaths","Spain Deaths"),
       col=c("blue","red"),
       lty = c(1, 1),
       cex = 0.6)
grid()

plot(EUMD$Group.1,EUMD$x, main="EU Gang of 5 Daily Deaths(mean and Sun",type="l",
     xlab="Date Reported",
     ylab="Daily Deaths",col="red")
lines(spain$dateRep,spain$deaths,col="blue",ylim=c(0,max(spain$deaths)))
legend("topleft",
       c("Mean Deaths","Spain Deaths"),
       col=c("blue","red"),
       lty = c(1, 1),
       cex = 0.6)
grid()
```

## Comparing Spain to UK Daily Cases and Deaths

```{r, echo=FALSE}
plot(spain$dateRep,spain$cases, main="Comparing Spain to UK Daily Cases Sun",type="l",
     xlab="Date Reported",
     ylab="Daily Cases",col="blue")
lines(UK$dateRep,UK$cases,col="red",ylim=c(0,max(spain$cases)))
legend("topleft",
       c("Spain Cases","UK Cases"),
       col=c("blue","red"),
       lty = c(1, 1),
       cex = 0.6)
grid()
```

```{r, echo=FALSE}
plot(spain$dateRep,spain$deaths, main="Comparing Spain to UK Daily Deaths Sun",type="l",
     xlab="Date Reported",
     ylab="Daily Deaths",
     ylim=c(0,max(spain$deaths)),col="blue")
lines(UK$dateRep,UK$deaths,col="red")
legend("topleft",
       c("Spain Deaths","UK Deaths"),
       col=c("blue","red"),
       lty = c(1, 1),
       cex = 0.6)
grid()
```

## Comparing Spain to Germany Daily Cases and Deaths

```{r, echo=FALSE}
plot(spain$dateRep,spain$cases, main="Comparing Spain to Germany Daily Cases Sun",type="l",
     xlab="Date Reported",
     ylab="Daily Cases",col="blue")
lines(DE$dateRep,DE$cases,col="red",ylim=c(0,max(spain$cases)))
legend("topleft",
       c("Spain Cases","Germany Cases"),
       col=c("blue","red"),
       lty = c(1, 1),
       cex = 0.6)
grid()
```

```{r, echo=FALSE}
plot(spain$dateRep,spain$deaths, 
     main="Comparing Spain to Germany Daily Deaths Sun",type="l",
     xlab="Date Reported",
     ylab="Daily Deaths",
     ylim=c(0,max(spain$deaths)),col="blue")
lines(DE$dateRep,DE$deaths,col="red")
legend("topleft",
       c("Spain Deaths","UK Deaths"),
       col=c("blue","red"),
       lty = c(1, 1),
       cex = 0.6)
grid()
```

## Comparing Spain to France Daily Cases and Deaths


```{r, echo=FALSE}
plot(spain$dateRep,spain$cases, main="Comparing Spain to France Daily Cases Sun",type="l",
     xlab="Date Reported",
     ylab="Daily Cases",col="blue")
lines(FR$dateRep,FR$cases,col="red",ylim=c(0,max(spain$cases)))
legend("topleft",
       c("Spain Cases","France Cases"),
       col=c("blue","red"),
       lty = c(1, 1),
       cex = 0.6)
grid()
```

```{r, echo=FALSE}
plot(spain$dateRep,spain$deaths, main="Comparing Spain to France Daily Deaths Sun",type="l",
     xlab="Date Reported",
     ylab="Daily Deaths",
     ylim=c(0,max(spain$deaths)),col="blue")
lines(UK$dateRep,UK$deaths,col="red")
legend("topleft",
       c("Spain Deaths","UK Deaths"),
       col=c("blue","red"),
       lty = c(1, 1),
       cex = 0.6)
grid()
```

## Comparing Spain to Italy Daily Cases and Deaths

```{r, echo=FALSE}
plot(spain$dateRep,spain$cases, main="Comparing Spain to Italy Daily Cases Sun",type="l",
     xlab="Date Reported",
     ylab="Daily Cases",col="blue")
lines(IT$dateRep,IT$cases,col="red",ylim=c(0,max(spain$cases)))
legend("topleft",
       c("Spain Cases","Italy Cases"),
       col=c("blue","red"),
       lty = c(1, 1),
       cex = 0.6)
grid()
```

```{r, echo=FALSE}
plot(spain$dateRep,spain$deaths, main="Comparing Spain to Italy Daily Deaths Sun",type="l",
     xlab="Date Reported",
     ylab="Daily Deaths",
     ylim=c(0,max(spain$deaths)),col="blue")
lines(IT$dateRep,IT$deaths,col="red")
legend("topleft",
       c("Spain Deaths","Italy Deaths"),
       col=c("blue","red"),
       lty = c(1, 1),
       cex = 0.6)
grid()
```

## Comparing Spain to Poland Daily Cases and Deaths

```{r, echo=FALSE}
plot(spain$dateRep,spain$cases, main="Comparing Spain to Poland Daily Cases Sun",type="l",
     xlab="Date Reported",
     ylab="Daily Cases",col="blue")
lines(PL$dateRep,PL$cases,col="red",ylim=c(0,max(spain$cases)))
legend("topleft",
       c("Spain Cases","Poland Cases"),
       col=c("blue","red"),
       lty = c(1, 1),
       cex = 0.6)
grid()
```

```{r, echo=FALSE}
plot(spain$dateRep,spain$deaths, main="Comparing Spain to Poland Daily Deaths Sun",type="l",
     xlab="Date Reported",
     ylab="Daily Deaths",
     ylim=c(0,max(spain$deaths)),col="blue")
lines(UK$dateRep,UK$deaths,col="red")
legend("topleft",
       c("Spain Deaths","UK Deaths"),
       col=c("blue","red"),
       lty = c(1, 1),
       cex = 0.6)
grid()
```